A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-06-09 main.tex
Improve Introduction of Multistage Programming.

Multistage programming enables us to generate and run fragments of code at runtime.
MetaOCaml\cite{oleg2014} provide features for multistage programming, which are quote, unquote, and running code.
Quote, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Unquote, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, unquote is supposed to use in a quoted fragment of code, that is, unquote cannot be used as an alternative of run.
For example, the following MetaOCaml expression

@@

Cross-stage Persistence (CSP) is another primitive of multistage programming, which enable us to embed computed values into a code value.
This is an example of CSP.
\begin{verbatim}
        let plusone = fun x -> x + 1 in
        let a = plusone 2 in  .< a >.
\end{verbatim}
This program evaluates to \verb|.< 3 >.| not  \verb|.< plusone 2 >.|.
This is because the variable \verb|a| is introduced into a code fragment using CSP.
Therefore, the variable \verb|a| is embedded after \verb|a| is calculated.
CSP is very important when we write practical programs 
because CSP enables us to use library functions in code fragments.

@@

The main application of multistage calculi is program optimization.
A famous example is \verb|power| function.
\verb|power| takes two arguments, which are base and exponent.
We can make specialized \verb|power| functions for given exponents with multistage calculi and optimize them by unrolling a loop in functions for given exponents.
\red{powerの実際のコードを持ってくるか?}

@@

Although unrolled \verb|add| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.

>>>>

Multistage programming enables us to generate and run fragments of code at runtime.
MetaOCaml\cite{oleg2014} provide features for multistage programming, which are brackets, escape, and run.
Brackets, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Escape, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, escape is supposed to use in a fragment of code, that is, escape cannot be used as an alternative of run.
For example, the following MetaOCaml expression

@@

Cross-stage Persistence (CSP) is another primitive of multistage programming, which enable us to embed computed values into a code value.
This is an example of CSP.
\begin{verbatim}
        let plusone = fun x -> x + 1 in
        let y = plusone 2 in  .< y >.
\end{verbatim}
This program evaluates to \verb|.< 3 >.| not  \verb|.< plusone 2 >.|.
This is because the variable \verb|y| is introduced into a code fragment using CSP.
Therefore, the variable \verb|y| is embedded after it was calculated.
CSP is very important when we write practical programs 
because CSP enables us to use library functions in code fragments as in \verb|.<List.combine [1;2] ['a';'b']>.|.

@@

The main application of multistage calculi is program optimization.
A famous example is power functions.
Usual power functions take two arguments, which are the base and the exponent.
We can make a specialized power function for given exponents with multistage programming and 
optimize them by unrolling a loop in functions for given exponents.
The optimized power function for the exponent of \verb|3| looks like \verb|power3 = fun x -> x * x * x|.
\verb|power3| is faster than ordinary \verb|power| function because it contains no loop.
Multistage programming can optimize functions which take more than two arguments 
by generating a code fragment optimized to a given argument.

@@

However, there is a sever problem in functions which are optimized by multistage programming.
Although unrolled \verb|add| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.
